### 2021-01-25 paper/arm.tex
Grrr

\begin{definition}
  \label{def:DS}
  % Let $[\FALSE/\D]$ be the substitution that performs $[\FALSE/\Dx{\aLoc}]$
  % for every $\aLoc$.
  Let $[\aForm/\Dx{*}]$ substitute $\aForm$ for every $\Dx{\bLoc}$.\\
  Let formula $\DL{\aLoc}{\amode}$ and substitution $\DS{\aLoc}{\amode}$ be defined:
  \begin{align*}
    \DL{\aLoc}{\mRLX}&=\TRUE
    &\DS{\aLoc}{\mRLX}&=[\TRUE/\Dx{\aLoc}] 
    \\
    \DL{\aLoc}{\mRA}&=\Dx{\aLoc}%\land\D{}
    &\DS{\aLoc}{\mRA}&=[\FALSE/\D]
    \\
    \DL{\aLoc}{\mSC}&=\Dx{\aLoc}%\land\D{}
    &\DS{\aLoc}{\mSC}&=[\FALSE/\D]
  \end{align*}
\end{definition}
\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$]
  \label{def:pomsets-down}
  Update \refdef{def:pomsets-ra} to (\ref{L4} unchanged):
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $\bForm[(\Qw{\aLoc}\land\aExp{=}\aVal)/\Qw{\aLoc}]\DS{\aLoc}{\amode}$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\bForm[\FALSE/\QS{\aLoc}{\amode}]\DS{\aLoc}{\amode}$,
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\DL{\aLoc}{\amode}\land \bForm[\FALSE/\QL{\aLoc}{\amode}]$.
  \end{enumerate}
\end{definition}
Load actions that require downgrading introduce $\Dx{\aLoc}$.
Relaxed stores on $\aLoc$ substitute true for $\Dx{\aLoc}$, whereas
synchronizing stores substitute false for $\Dx{\aLoc}$.

@@

\begin{example}
  Revisiting \refex{ex:rrd1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \event{b}{\DR{y}{1}}{}
          \xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{r\mid\DW{x}{0}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing sequentially:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \SEMI
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{\RW\limplies r\mid\DW{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Embedding the thread in thread group, we have, as desired:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \SEMI
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{\FALSE\limplies r\mid\DW{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
\end{example}

@@

\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$/$\xRRD$]
  \label{def:pomsets-arm}
  Update \refdef{def:pomsets-trans} to (\ref{L4} unchanged):
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm\noSUB{[\aExp/\aLoc]}$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm\noSUB{[\aExp/\aLoc]}$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    % $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
    % \item[\ref{L6})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode}{}
    \land (\RW\limplies\bForm)$.
  \end{enumerate}  
\end{definition}

Every \armeight{} execution is allowed by \refdef{def:pomsets-arm}.  The
proof of this fact is simplified by the recent characterization of
\armeight{} in terms of \EGC{} \cite[\textsection
B2.3.6]{arm-reference-manual}.  Under \EGC{}, an \armeight{} execution is a
linearization of per-location program order and a subset of local-order.
Every such linearization is also a valid pomset under
\refdef{def:pomsets-arm}.

>>>>

\begin{definition}
  \label{def:DS}
  % Let $[\FALSE/\D]$ be the substitution that performs $[\FALSE/\Dx{\aLoc}]$
  % for every $\aLoc$.
  Let $[\aForm/\Dx{*}]$ substitute $\aForm$ for every $\Dx{\bLoc}$.\\
  Let formula $\DL{\aLoc}{\amode}$ and substitution $\DS{\aLoc}{\amode}$ be defined:
  \begin{align*}
    \DL{\aLoc}{\mRLX}&=\TRUE
    &\DS{\aLoc}{\mRLX}&=[\TRUE/\Dx{\aLoc}] 
    \\
    \DL{\aLoc}{\mRA}&=\Dx{\aLoc}%\land\D{}
    &\DS{\aLoc}{\mRA}&=[\FALSE/\D]
    \\
    \DL{\aLoc}{\mSC}&=\Dx{\aLoc}%\land\D{}
    &\DS{\aLoc}{\mSC}&=[\FALSE/\D]
  \end{align*}
\end{definition}
\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$]
  \label{def:pomsets-down}
  Update \refdef{def:pomsets-ra} to (\ref{L4} unchanged):
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies
    \begin{math}
      \bForm
      \DS{\aLoc}{\amode}
      [(\Qw{\aLoc}\land\aExp{=}\aVal)/\Qw{\aLoc}],
    \end{math}
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      \bForm
      \DS{\aLoc}{\amode}
      [\FALSE/\QS{\aLoc}{\amode}],
    \end{math}
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies
    \begin{math}
      \aVal{=}\aReg\limplies\bForm,
    \end{math}
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      \DL{\aLoc}{\amode}\land \bForm[\FALSE/\QL{\aLoc}{\amode}].
    \end{math}
  \end{enumerate}
\end{definition}
Load actions that require downgrading introduce $\Dx{\aLoc}$.
Relaxed stores on $\aLoc$ substitute true for $\Dx{\aLoc}$, whereas
synchronizing stores substitute false for $\Dx{\aLoc}$.

@@

\begin{example}
  Revisiting \refex{ex:rrd1} and eliding irrelevant transformers:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \event{b}{\DR{y}{1}}{}
          \xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{r\mid\DW{x}{0}}{}
          % \xform{x1d}{\bForm[\TRUE/\Dx{\aLoc}]}{below=of c}
          % \xform{x1i}{\lnot\Qw{y}\land\bForm[\TRUE/\Dx{y}]}{below=of x1d}
          % \xo{c}{x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Composing sequentially:
  \begin{align*}
    \begin{gathered}[t]
      \PR{y}{r}
      \SEMI
      \IF{r}\THEN s\GETS x\FI
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{\RW\limplies r\mid\DW{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Embedding the thread in thread group, \ref{T3} yields:
  \begin{align*}
    \begin{gathered}[t]
      % \PR{y}{r}
      % \SEMI
      % \IF{r}\THEN s\GETS x\FI
      % \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{b}{\DR{y}{1}}{}
          \event{c}{\FALSE\limplies r\mid\DW{x}{0}}{right=of b}
          %\xform{x1i}{\RW\limplies\bForm}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  The precondition of $\DWP{x}{0}$ is a tautology, as required.
\end{example}

@@

\begin{definition}[$\xCO$/$\xRASC$/$\xDGR$/$\xRRD$]
  \label{def:pomsets-arm}
  Update \refdef{def:pomsets-trans} to (\ref{L4} unchanged):
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies
    \begin{math}
      %(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land
      \bForm %\noSUB{[\aExp/\aLoc]},
      [(\Qw{\aLoc}\land\aExp{=}\aVal)/\Qw{\aLoc}]
      \DS{\aLoc}{\amode}
    \end{math}
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      %\lnot\Qw{\aLoc}\land 
      \bForm %\noSUB{[\aExp/\aLoc]}.
      [\FALSE/\QS{\aLoc}{\amode}]
      \DS{\aLoc}{\amode}
    \end{math}
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    % $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
    % \item[\ref{L6})]
    $\aTr{\cEvs}{\bForm}$ implies
    \begin{math}
      % \lnot\Qx{\aLoc} \land
      \DL{\aLoc}{\amode}{} \land (\RW\limplies\bForm[\FALSE/\QL{\aLoc}{\amode}]).
    \end{math}
  \end{enumerate}  
\end{definition}

\begin{example}
  $\xRRD$ does not adversely affect subscription (\refex{ex:subscription}).
  \begin{align*}
    \begin{gathered}[t]
      \PR[\mRA]{y}{r}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and .5em]
          \raevent{b}{\QL{y}{\mRA}\mid\DR[\mRA]{y}{1}}{}
          \xform{x1i}{\Dx{y}\land(\RW\limplies\bForm[\FALSE/\QL{\aLoc}{\mRA}])}{right=of b} %x1d}
        \end{tikzinline}}  
    \end{gathered}  
    &&
    \begin{gathered}[t]
      \PR{x}{s}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \event{c}{\QL{y}{\mRA}\mid\DW{x}{0}}{}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Since $\QL{y}{\mRA}[\FALSE/\QL{\aLoc}{\mRA}]$ is false, we have:
  \begin{align*}
    \begin{gathered}[t]
      \PR[\mRA]{y}{r}
      \SEMI
      \PR{x}{s}
      \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\QL{y}{\mRA}\mid\DR[\mRA]{y}{1}}{}
          \event{c}{\Dx{y}\land(\RW\limplies\FALSE)\mid\DW{x}{0}}{right=of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  If this is the complete thread, we apply \ref{T3} to yield:
  \begin{align*}
    \begin{gathered}[t]
      % \PR[\mRA]{y}{r}
      % \SEMI
      % \PR{x}{s}
      % \\
      \hbox{\begin{tikzinline}[node distance=0.5em and 1.5em]
          \raevent{b}{\QL{y}{\mRA}\mid\DR[\mRA]{y}{1}}{}
          \event{c}{\Dx{y}\land(\FALSE\limplies\FALSE)\mid\DW{x}{0}}{right=of b}
        \end{tikzinline}}  
    \end{gathered}  
  \end{align*}
  Although $(\FALSE\limplies\FALSE)$ is a tautology, $\Dx{y}$ is not.  This
  pomset can only contribute to a top-level pomset if the thread is preceded
  by a relaxed write to $y$, allowing $\DRP[\mRA]{y}{1}$ to be downgraded to
  a relaxed read.
\end{example}

Every \armeight{} execution is allowed by \refdef{def:pomsets-arm}.  The
proof of this fact is simplified by the recent characterization of
\armeight{} in terms of \emph{external global completion} (\EGC{})
\cite[\textsection B2.3.6]{arm-reference-manual}.  Under \EGC{}, an
\armeight{} execution is a linearization of per-location program order and a
subset of local-order.  Every such linearization is also a valid pomset under
\refdef{def:pomsets-arm}.
